On exponential sums arising from the classical doubling method

### 2025-03-30 main.tex
minor adjustments in main theorem proof

\calvin{I think I would like to include a main theorem here even if it's not super precise just to let the reader know what we highlight what we about the sum}
\begin{theorem}[\Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig}, \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}]
	For a classical group $G$ and a multiplicative character $\chi$ as above, $\genHermitianJacobiKernel{\hermitianSpace}{\chi}$ are stable functions. We have that
	$$\posDblJacobiSumScalar{\pi}{\chi} =
		\varepsilon_{\algebraicGroup{H}} c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right),$$
		and that
	$$\negDblJacobiSumScalar{\pi}{\chi} = \begin{dcases}
		0 & $G$ \text{ is an odd special orthogonal group},\\
		\centralCharacter{\pi}\left(-1\right) \posDblJacobiSumScalar{\pi}{\chi} &  \text{otherwise.}
	\end{dcases}$$
Here $c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)$ is an explicit factor and
$g_T\left(\chi, \theta, \fieldCharacter\right)$ is a product of Gauss sums, see \Cref{subsec:normalization-factor} and \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}.
\end{theorem}
In this paper, we study the doubling method Jacobi sums $\dblJacobiSum{\pi}{\chi}$. We compute the scalars $\dblJacobiSumScalar{\pi}{\chi}$ in terms of the Deligne--Lusztig data corresponding to $\pi$ and show that, up to some well understood factors, they match the Kondo Gauss sum of the conjectural functorial transfer of $\pi$ to a general linear group. This is done by combining a couple of ingredients. The first ingredient is a lemma of Saito--Shinoda \cite{SaitoShinoda2000} that allows us to explicitly compute the analogous character sums for Deligne--Lusztig virtual representations $R_{T, \theta}$.  The second ingredient, which is a key ingredient for this computation, is that these Jacobi sums are well behaved, in the sense that they are constant on Lusztig series. To show this, we utilize a multiplicativity property of these Jacobi sums from \cite{GirschZelingher2025} and combine it with a careful analysis of the Deligne--Lusztig data corresponding to $\pi$ after functorial transfer.

@@

\calvin{end alternate writeup}
\end{comment}
\calvin{start alternate writeup 2 of section 3.4/3.4.1}
\subsection{Stability (one option)}
The goal of this section is to compute $\dblJacobiSumScalar{\pi}{\fieldCharacter}$ proving \Cref{thm:Phi_stable}:
\begin{theorem}
\label{thm:doubling-method-gamma-factor-for-deligne-lusztig} For either $G=G$ or its extension $\GroupExtension{G}$. If $c_{\pi}$ is nonzero in the virtual representation $R_T(\theta) = \sum c_{\pi} \pi$ then
$$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = g_{T \cap G}\left(\chi,\theta|_{T \cap G}, \fieldCharacter\right)$$
\end{theorem}
\calvin{add in $\chi$} the right hand side of which we showed in \Cref{lem:indep_geo_conj} only depends on the geometric conjugacy class of $(T,\theta)$ hence on the Lusztig series $\LusztigSeries{G}{s}$ which contains $\pi$, implying \Cref{thm:Phi_stable}.

@@

Our proof will be via induction on $\dim_{\quadraticExtension} \hermitianSpace$. First we will use \Cref{thm:multiplicativity-in-terms-of-gamma-factors}  and MULTIPLICAATIVITY-OF-GTTHETA to prove \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} for any non cuspidal $\pi$ assuming \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} for all Levi factors. \\

Combining this with our inductive hypothesis implies \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} for non cupsidal $\pi$.
By \Cref{lem:indep_geo_conj}, we know that for $R_T(\theta) = \sum c_{\pi} \pi$, 
$$\sum c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = (\sum c_{\pi})g_{T \cap G}\left(\theta|_{T \cap G}, \fieldCharacter\right)$$. Subtracting our the equality for non-cupsidal $\pi$ we are left with:
$$ \sum_{\pi \textnormal{ cuspidal}} c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \left(\sum_{\pi \textnormal{ cuspidal}} c_{\pi}\right)g_{T \cap G}\left(\theta|_{T \cap G}, \fieldCharacter\right)$$.

@@

So the proof of \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} is complete.

\calvin{end alternate writeup 2}

>>>>

\calvin{I think I would like to include a main theorem here even if it's not super precise just to let the reader know what we highlight what we about the sum}
\begin{theorem}[\Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig}, \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}]
	For a classical group $G$ and a multiplicative character $\chi$ as above, $\genHermitianJacobiKernel{\hermitianSpace}{\chi}$ are stable functions. We have that
	$$\posDblJacobiSumScalar{\pi}{\chi} =
		\varepsilon_{\algebraicGroup{H}} c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right),$$
		and that
	$$\negDblJacobiSumScalar{\pi}{\chi} = \begin{dcases} 0 & G \text{ is an odd special orthogonal group},\\
		\centralCharacter{\pi}\left(-1\right) \posDblJacobiSumScalar{\pi}{\chi} &  \text{otherwise.}
	\end{dcases}$$
Here $c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)$ is an explicit factor and
$g_T\left(\chi, \theta, \fieldCharacter\right)$ is a product of Gauss sums, see \Cref{subsec:normalization-factor} and \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}.
\end{theorem}
In this paper, we study the doubling method Jacobi sums $\dblJacobiSum{\pi}{\chi}$. We compute the scalars $\dblJacobiSumScalar{\pi}{\chi}$ in terms of the Deligne--Lusztig data corresponding to $\pi$ and show that, up to some well understood factors, they match the Kondo Gauss sum of the conjectural functorial transfer of $\pi$ to a general linear group. This is done by combining a couple of ingredients. The first ingredient is a lemma of Saito--Shinoda \cite{SaitoShinoda2000} that allows us to explicitly compute the analogous character sums for Deligne--Lusztig virtual representations $R_{T, \theta}$.  The second ingredient, which is a key ingredient for this computation, is that these Jacobi sums are well behaved, in the sense that they are constant on Lusztig series. To show this, we utilize a multiplicativity property of these Jacobi sums from \cite{GirschZelingher2025} and combine it with a careful analysis of the Deligne--Lusztig data corresponding to $\pi$ after functorial transfer.

@@

\calvin{end alternate writeup}
\end{comment}
\calvin{start alternate writeup 2 of section 3.4/3.4.1}
\subsection{Stability (one option)}\label{subsec:stability}
The goal of this section is to compute $\dblJacobiSumScalar{\pi}{\fieldCharacter}$ proving \Cref{thm:Phi_stable}:
\begin{theorem}
\label{thm:doubling-method-gamma-factor-for-deligne-lusztig} For either $G=G$ or its extension $\GroupExtension{G}$. If $c_{\pi}$ is nonzero in the virtual representation $R_T(\theta) = \sum c_{\pi} \pi$ then
$$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = g_{T \cap G}\left(\chi,\theta|_{T \cap G}, \fieldCharacter\right)$$
\end{theorem}
\calvin{add in $\chi$} the right hand side of which we showed in \Cref{lem:indep_geo_conj} only depends on the geometric conjugacy class of $(T,\theta)$ hence on the Lusztig series $\LusztigSeries{G}{s}$ which contains $\pi$, implying \Cref{thm:Phi_stable}.

@@

Our proof will be via induction as descriped in \ref{subsec:stability}.
%First we will use \Cref{thm:multiplicativity-in-terms-of-gamma-factors}  and MULTIPLICAATIVITY-OF-GTTHETA to prove \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} for any non cuspidal $\pi$ assuming \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} for all Levi factors. \\
\begin{proof}[Proof of (A)]
\calvin{should be short proof. Might not even need to put--just need to say both are multiplicative}
\end{proof}
\begin{proof}[Proof of (B)]

%Combining this with our inductive hypothesis implies \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} for non cupsidal $\pi$.
By \Cref{lem:indep_geo_conj}, we know that for $R_T(\theta) = \sum c_{\pi} \pi$, 
$$\sum c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = (\sum c_{\pi})g_{T \cap G}\left(\theta|_{T \cap G}, \fieldCharacter\right).$$Subtracting our the equality for non-cupsidal $\pi$ we are left with:
$$ \sum_{\pi \textnormal{ cuspidal}} c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \left(\sum_{\pi \textnormal{ cuspidal}} c_{\pi}\right)g_{T \cap G}\left(\theta|_{T \cap G}, \fieldCharacter\right)$$.

@@

So the proof of \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} is complete.
\end{proof}
\calvin{end alternate writeup 2}
